---
format:
  html:
    toc: true
    toc-depth: 4
    page-layout: full
---

# Analysis

Alix: "Our goal is to define the top 20% of all Hubbard Hall donors from the past 11 years (I think that was how far back we said this went) and then also from the past 5 years.

We are looking for contact family name, total giving all time, last gift date, last FY total giving amount, and for the top 20% a breakdown of what their gift categories were (we want a complete donor profile, so for each donor we want to see the total contribution amount that includes donations, class purchases, ticket buys, etc.)

We have a couple of questions for Judy additionally, that may be answered after Don has a go. Namely, we are looking for where Sue's donations from Opus live, and also where Kathy Rhoome's donations live to make sure that donations like these are included in the total.  Also, we want to find out if there is a way to link businesses who give to individuals who exist in the database as a soft credit?  This doesn't apply to all businesses, but for example, is MSK linked to Jason in a way where you can pull total giving for an individual and see his corporate contributions as well?"

Top 100 donors all time, 
and top 100 donors last 5 years

Also, seeing if there is a way to include corporate contributions for any of these individuals as appropriate.

From Judy: "I tested whether a deceased household member’s donations were still included in the household total, and they are. If the primary account is still active and the other household members data still exists then it is included. If the other household members account is deleted, then the data is deleted."


## Setup

```{r}
#| label: setup

source(here::here("libraries.R"))

hhdr <- here::here("data-raw")
neon <- fs::path(hhdr, "neon")
judy <- fs::path(hhdr, "judy")
rds <- fs::path(hhdr, "rds")
work <- fs::path(hhdr, "work")

```

## Retrieve prepared data

```{r}
#| label: get-data

linkages <- readRDS(fs::path(rds, "linkages.rds"))
accounts <- readRDS(fs::path(rds, "accounts.rds"))
contacts <- readRDS(fs::path(rds, "contacts.rds"))
donations <- readRDS(fs::path(rds, "donations.rds"))
events <- readRDS(fs::path(rds, "events.rds"))
registrations <- readRDS(fs::path(rds, "registrations.rds"))

xljudy <- read_excel(
  fs::path(judy, "HH Donor List - All Donations $15+.xlsx"),
  sheet = "List by Total Donations"
)

```

## Analysis

Our goal is to define the top 20% of all Hubbard Hall donors from the past 11 years (I think that was how far back we said this went) and then also from the past 5 years.

We are looking for contact family name, total giving all time, last gift date, last FY total giving amount, and for the top 20% a breakdown of what their gift categories were (we want a complete donor profile, so for each donor we want to see the total contribution amount that includes donations, class purchases, ticket buys, etc.)

[Dashboard](https://hubbardhall.app.neoncrm.com/np/admin/content/contentList.do)


```{r}
#| label: account-linkages

acctlink <- accounts |>
  # I DON'T YET KNOW WHY I HAVE TO ADJUST Cora Howe -- her partner Robert is listed as primary in the data
  mutate(
    contact_type = case_when(
      hhid == "54" & accountid == "34713" ~ "Primary Household Contact", # Cora May Howe
      hhid == "54" & accountid == "34712" ~ "Household Contact", # Robert Howe, deceased and DNC
      .default = contact_type
    )
  ) |>
  # filter(account_type=="Individual", group=="account", id=="38532")
  filter(account_type == "Individual") |>
  filter(
    group == "account" |
      (group == "household" &
        str_detect(contact_type, "Primary Household Contact"))
  ) |>
  # captures PHC in combination with anything else CAUTION: keeps company contacts
  select(
    id,
    name,
    group,
    primary_contact = fullname,
    contact_type,
    deceased,
    donotcontact
  )

# force hhid keeps
force_keeps <- c(
  "403", # Bloomington
  "530", # Brand
  "1620" # Aldrich
)

# special hhid drops related to Judy
force_drops <- c("224") # Caroline Eastman

```

## Donors, all time

```{r}
#| label: alix-table1

# get hh by fiscal year
don1 <- donations |>
  filter(account_type == "Individual") |>
  filter(
    donation_type %in% c("DONATION", "PLEDGEPAYMENT"),
    donation_status == "SUCCEEDED"
  ) |>
  filter(donation_date <= "2025-06-30") |>
  mutate(
    hhfy = ifelse(
      month(donation_date) >= 7,
      year(donation_date) + 1,
      year(donation_date)
    )
  )

# get summaries for each household across ALL donations
ns(don1)
skim(don1) # no missing donations

idvars <- c("id", "name", "group", "hhid", "hhname", "accountid")

hhalldon <- don1 |>
  group_by(id, name) |>
  mutate(
    donation_all = sum(donation),
    donation_last5years = sum(donation[hhfy %in% 2021:2025]),
    donation_max = max(donation),
    n_max_donation = sum(donation == donation_max, na.rm = TRUE),
    date_of_max = donation_date[donation == donation_max][1],
    campaign_of_max = campaign[donation == donation_max][1],
    last_donation_date = max(donation_date),
    fy_of_last_donation = hhfy[donation_date == last_donation_date][1],
    total_donation_latest_donation_fy = sum(donation[
      hhfy == fy_of_last_donation
    ])
  ) |>
  slice_head(n = 1) |>
  ungroup() |>
  arrange(desc(donation_all), id) |>
  select(
    any_of(idvars),
    donation_all,
    donation_last5years,
    donation_max,
    n_max_donation,
    date_of_max,
    campaign_of_max,
    last_donation_date,
    fy_of_last_donation,
    total_donation_latest_donation_fy
  )

```


```{r}

## Bring in additional info
hhall_enhanced <- hhalldon |>
  left_join(acctlink |> select(-c(group)), by = join_by(id, name)) |>
  left_join(
    contacts |> select(id, name, pfname, plname, address1:pphone),
    by = join_by(id, name)
  ) |>
  filter(!deceased, accountid != "36805") |>
  select(-deceased)

ns(hhall_enhanced)

```


## Donors, 2021+


```{r}
#| label: alix-table2

# get hh by fiscal year
don1_2021plus <- don1 |>
  filter(account_type == "Individual") |>
  mutate(donation_all = sum(donation), .by = c(id, name)) |>
  filter(donation_date >= "2020-07-01") |>
  filter(
    donation_type %in% c("DONATION", "PLEDGEPAYMENT"),
    donation_status == "SUCCEEDED"
  ) |>
  mutate(
    hhfy = ifelse(
      month(donation_date) >= 7,
      year(donation_date) + 1,
      year(donation_date)
    )
  )

# get summaries for each household across ALL donations
ns(don1_2021plus)
skim(don1_2021plus) # no missing donations

idvars <- c("id", "name", "group", "hhid", "hhname", "accountid")

hhalldon_2021plus <- don1_2021plus |>
  group_by(id, name) |>
  mutate(
    donation_last5years = sum(donation[hhfy %in% 2021:2025]),
    donation_max = max(donation),
    n_max_donation = sum(donation == donation_max, na.rm = TRUE),
    date_of_max = donation_date[donation == donation_max][1],
    campaign_of_max = campaign[donation == donation_max][1],
    last_donation_date = max(donation_date),
    fy_of_last_donation = hhfy[donation_date == last_donation_date][1],
    total_donation_latest_donation_fy = sum(donation[
      hhfy == fy_of_last_donation
    ])
  ) |>
  slice_head(n = 1) |>
  ungroup() |>
  arrange(desc(donation_last5years), id) |>
  select(
    any_of(idvars),
    donation_all,
    donation_last5years,
    donation_max,
    n_max_donation,
    date_of_max,
    campaign_of_max,
    last_donation_date,
    fy_of_last_donation,
    total_donation_latest_donation_fy
  )

```


```{r}

## Bring in additional info
hhall_enhanced_2021plus <- hhalldon_2021plus |>
  left_join(acctlink |> select(-c(group)), by = join_by(id, name)) |>
  left_join(
    contacts |> select(id, name, pfname, plname, address1:pphone),
    by = join_by(id, name)
  ) |>
  filter(!deceased, accountid != "36805") |>
  select(-deceased)

ns(hhall_enhanced_2021plus)

```

## Event types

```{r}
#| label: event-categorizations

regorder <- c("performance", "class", "community", "other")

events_data <- events |>
  select(
    event_majorcat,
    event_category,
    event_name,
    event_start_date,
    eventid,
    event_code
  ) |>
  mutate(event_majorcat = factor(event_majorcat, levels = regorder)) |>
  arrange(event_majorcat, event_category, eventid, event_name)

range_write(
  ss,
  events_data,
  sheet = "event_categorizations",
  range = "A13",
  col_names = TRUE,
  reformat = FALSE
)


```

## First cut registrations

```{r}

skim(registrations)
ns(registrations)
count(registrations, type)
count(registrations, event_status)

registrations |>
  filter(
    type %in% c("Registrant", "Registrant & Attendee"),
    event_status == "SUCCEEDED"
  ) |>
  pull(amount) |>
  sum()

reg1 <- registrations |>
  filter(!is.na(id)) |>
  filter(
    type %in% c("Registrant", "Registrant & Attendee"),
    event_status == "SUCCEEDED"
  ) |>
  summarise(amount = sum(amount), .by = c(id, name, group, event_majorcat))
sum(reg1$amount)

donor_events <- hhall_enhanced |>
  filter(!donotcontact) |>
  select(id, name, group, donation_all) |>
  left_join(reg1 |> select(-group), by = join_by(id, name)) |>
  filter(!is.na(event_majorcat)) |>
  mutate(event_majorcat = factor(event_majorcat, levels = regorder)) |>
  mutate(event_sum = sum(amount, na.rm = TRUE), .by = c(id, name)) |>
  arrange(event_majorcat) |>
  pivot_wider(
    names_from = event_majorcat,
    values_from = amount,
    values_fill = 0
  ) |>
  arrange(desc(donation_all)) |>
  mutate(event_pct = event_sum / donation_all) |>
  relocate(event_pct, .before = event_sum) |>
  filter(row_number() <= 100)


# Registrants" or "Registrant & Attendees"

```

## Winnow and write to google sheets

```{r}

gs4_auth() # always
# gswb <- gs4_create("donors") # ONETIME
# sheet_write(donfinal, gswb, sheet = "all_donors") # writing a full sheet

ssid <- "1xwNYOddcJSRmAvaR9DlT1Aixrs1WciwVYUvv3AzmdM0"
donors_workbook <- "https://docs.google.com/spreadsheets/d/1xwNYOddcJSRmAvaR9DlT1Aixrs1WciwVYUvv3AzmdM0"
(ss <- gs4_get(donors_workbook))
ss

# sheet_add(ss, sheet = NULL, ..., .before = NULL, .after = NULL)

```


```{r}
#| label: do-not-contact

glimpse(hhall_enhanced)

dnc <- hhall_enhanced |>
  filter(donotcontact, donation_all >= 1000) |>
  mutate(explanation = "") |>
  relocate(explanation, .after = hhname)

# sheet_add(ss, sheet = "do_not_contact")

# do NOT accidentally overwrite do_not_contact because it has hand-entered explanations
# range_write(
#   ss,
#   dnc,
#   sheet = "do_not_contact",
#   range = "A10",
#   col_names = TRUE,
#   reformat = FALSE
# )

```


```{r}
#| label: top100-all-time

glimpse(hhall_enhanced)

top100 <- hhall_enhanced |>
  filter(!donotcontact) |>
  arrange(desc(donation_all))

# sheet_add(ss, sheet = "top100_all_time")

range_write(
  ss,
  top100,
  sheet = "top100_all_time",
  range = "A10",
  col_names = TRUE,
  reformat = FALSE
)


```


```{r}
#| label: top100-last5years

top100_last_5years <- hhall_enhanced_2021plus |>
  filter(!donotcontact) |>
  arrange(desc(donation_last5years))

# sheet_add(ss, sheet = "top100_all_time")

range_write(
  ss,
  top100_last_5years,
  sheet = "top100_last_5years",
  range = "A10",
  col_names = TRUE,
  reformat = FALSE
)

```



```{r}
#| label: donor-events

# sheet_add(ss, sheet = "donor_events")

range_write(
  ss,
  donor_events,
  sheet = "donor_events",
  range = "A10",
  col_names = TRUE,
  reformat = FALSE
)

```


## Repeat for last 5 years

```{r}

sum(donor_events$event_sum)

```



## Quick plot: Household donor spending and performances, 2021 plus

```{r}
#| label: donations-events
#| eval: false

glimpse(donations)
glimpse(registrations)

cashcc <- donations |> filter(accountid == "36805")
sum(cashcc$donation) # 26317.43

don1 <- donations |>
  rename(account_name = fullname) |>
  filter(account_type == "Individual", year(donation_date) >= 2021) |>
  filter(
    !str_detect(account_name, "Cash and Credit Card Miscellaneous Transactions")
  ) |>
  summarise(
    n = n(),
    donation = sum(donation, na.rm = TRUE),
    .by = c(id, name, group)
  ) |>
  filter(donation >= 15) |>
  arrange(desc(donation))

regs1 <- registrations |>
  filter(year(event_start_date) >= 2021) |>
  filter(str_detect(type, "Registrant"), event_status == "SUCCEEDED") |>
  select(id, name, group, hhid, hhname, accountid, account_name, amount)
skim(regs1)

regs2 <- regs1 |>
  summarise(amount = sum(amount, na.rm = TRUE), .by = c(id, name, group)) |>
  arrange(desc(amount))

#  acccount id	name 36805	Cash and Credit Card Miscellaneous Transactions in 2021 plus was 28 transactions, $3890, all donotcontact

regs1 |> filter(is.na(id))

comp <- don1 |>
  left_join(regs1, by = join_by(hhid, hhname)) |>
  mutate(amount = replace_na(amount, 0))
skim(comp)

comp |>
  ggplot(aes(donation, amount)) +
  geom_point()

comp |>
  ggplot(aes(amount, donation)) +
  geom_point()

comp |>
  filter(donation >= 5000, donation < 20000, amount < 2000) |>
  ggplot(aes(amount, donation)) +
  geom_point()

```


## Investigations to determine how donations should be counted

I compare the items below to Neon query (report) results.

```{r}

names(xljudy)


df <- vroom(
  fs::path(neon, "HH Donor List - All Donations $15+.csv"),
  col_types = cols(.default = col_character())
)
ns(df)

count(donations, account_type)
count(donations, donation_type)
count(donations, payment_status)
count(donations, individual_type)

# filter account level data
don1 <- donations |>
  filter(account_type == "Individual") |>
  # filter(!deceased, !donotcontact) |> # these pertain to the individual account
  filter(
    donation_type %in% c("DONATION", "PLEDGEPAYMENT"),
    donation_status == "SUCCEEDED"
  ) |>
  # special drops related to Judy
  filter(
    !hhid %in%
      c("224")
  ) # Caroline Eastman

# summarise to household level
don2 <- don1 |>
  summarise(donation = sum(donation), .by = c(id, name)) |>
  arrange(desc(donation))

# link to accounts to see if primary householder is deceased or do not contact
acctlink <- accounts |>
  # I DON'T YET KNOW WHY I HAVE TO ADJUST Cora Howe -- her partner Robert is listed as primary in the data
  mutate(
    contact_type = case_when(
      hhid == "54" & accountid == "34713" ~ "Primary Household Contact", # Cora May Howe
      hhid == "54" & accountid == "34712" ~ "Household Contact", # Robert Howe, deceased and DNC
      .default = contact_type
    )
  ) |>
  # filter(account_type=="Individual", group=="account", id=="38532")
  filter(account_type == "Individual") |>
  filter(
    group == "account" |
      (group == "household" &
        str_detect(contact_type, "Primary Household Contact"))
  ) |>
  # captures PHC in combination with anything else CAUTION: keeps company contacts
  select(
    id,
    name,
    group,
    primary_contact = fullname,
    contact_type,
    deceased,
    donotcontact
  )

acctlink |> filter(id == "38532")
acctlink |> filter(str_detect(name, "Brosch"))
donations |> filter(str_detect(name, "Brosch"))

acctlink |> summarise(n = n(), .by = c(id, name)) |> count(n)


don3 <- don2 |>
  left_join(acctlink, by = join_by(id, name))

ddnc <- don3 |>
  filter(deceased | donotcontact)
anyDuplicated(ddnc$id)


force_keeps <- c(
  "403", # Bloomington
  "530", # Brand
  "1620" # Aldrich
)
don4 <- don3 |>
  filter(!(id %in% ddnc$id & !id %in% force_keeps)) |>
  select(-deceased, -donotcontact) |>
  filter(donation >= 15, id != "36805") |>
  arrange(desc(donation), as.numeric(id))

skim(don4)
don4 |> filter(is.na(group))
don4 |> filter(id == "1620")

don4 |> writexl::write_xlsx(fs::path(work, "don4.xlsx"))


donations |>
  filter(id == "21") |>
  select(id, name, donation, donation_date, payment_status) |>
  janitor::adorn_totals()
donations |>
  filter(id == "21") |>
  select(id, name, donation, donation_date, payment_status) |>
  arrange(donation_date) |>
  janitor::adorn_totals()
# this explains it: 21 Wendy Duvall   100.00    2018-09-18       CANCELED

```
